---
title: "Practical Machine Learning"
date: "Saturday, January 24, 2015"
output: html_document
---

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).


For this project, we analyzed the provided data to determine what activity an individual perform. To do this we made use of many libraries, this allowed us to generate correct answers for each of the 20 test data cases provided in this assignment. We made use of a seed value for consistent results.

```{r}
library(caret)
library(RANN)
library("randomForest")
set.seed(2015)
```

We load the data.

```{r,cache=TRUE}
training <- read.csv("training.csv")
testing <- read.csv("testing.csv")
training$classe <- as.factor(training$classe) 
summary(training)
str(training)
```

We clean the variables.

```{r}
NAindex <- apply(training,2,function(x) {sum(is.na(x))}) 
training <- training[,which(NAindex ==0)]
testing <- testing[,which(NAindex ==0)]
```
We preprocess the variables

```{r,cache=TRUE}
v <- which(lapply(training, class) %in% "numeric")
preObj <-preProcess(training[,v],method=c('knnImpute', 'center', 'scale'))
trainLess1 <- predict(preObj, training[,v])
trainLess1$classe <- training$classe
testLess1 <-predict(preObj,testing[,v])
```
We remove the values near to zero.

```{r}
nzv <- nearZeroVar(trainLess1,saveMetrics=TRUE)
trainLess1 <- trainLess1[,nzv$nzv==FALSE]
nzv <- nearZeroVar(testLess1,saveMetrics=TRUE)
testLess1 <- testLess1[,nzv$nzv==FALSE]
```


The training set is divided in two parts, one for training and the other for cross validation

```{r}
inTrain = createDataPartition(trainLess1$classe, p = .75, list=FALSE)
training = trainLess1[inTrain,]
crossValidation = trainLess1[-inTrain,]
dim(training)
```


The algorithm that we choose is random forest because, first we don't know the interaction between the variables, the random forest is built in cross-validation component that gives an unbiased estimation of the forest(OBB), the OBB can help to tune the forest, a random forest can be used to determine the importance of the variables, and the individual trees can be pulled out and be examined.

```{r,cache=TRUE}
modFit <-randomForest(classe ~., data=training )
modFit
```
We calculate the acurracy in the training and cross validation

```{r}
cvPred <- predict(modFit, crossValidation)
confusionMatrix(cvPred, crossValidation$classe)
```

Predictions on the real testing set

```{r}
testingPred <- predict(modFit, testLess1)
testingPred
```

